Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-06-08 code/manuscript.Rmd
Updated continous sampling ratio and manuscript

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events including static changes and gradual changes denoted as operational drift occurs during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability. We also assessed the impact of dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.

@@

# colors
ls_colors <- c("Baseline" = "grey10",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#6baed6",
               "Randomized\n(50/50)" = "#6baed6",
               "Randomized\n(20/80)" = "#6baed6", 
               "Randomized\n(10/90)" = "#6baed6", 
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Dropped (24 months)" = "#4292c6",
               "Kept (24 months)" = "#08519c",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we proposed a M&V method that
adopts the randomized crossover design [@raftery2024], a gold standard
from medical, agricultural and online controlled experimental (i.e. A/B
testing) studies [@blackston2019; @duan2013; @gupta2019; @mie2022;
@munro; @raseduzzaman2017]. An early application of this randomized
design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance on two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defined stopping criteria to end the M&V period early
after reaching a target level of uncertainty, among other improvements,
such as using blocked randomization by weekday. The full framework is
detailed in a previously published study [@raftery2024] along with a
case study application to a real building, and other example use cases.
In summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially shortly after the test
begins. In addition, once the desired estimate uncertainty target is
achieved, analysts can terminate the M&V study and switch to 100%
intervention. The key advantage of randomization is that if control
strategies are sampled with equal probability, the influence of other
confounding factors such as occupancy change (commonly known as one of
the non-routine events) and other more subtle long term changes in
building energy use are likely to be evenly distributed among
measurements, leading to a more accurate and unbiased assessment of the
intervention effect.

@@

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly use the metered electricity and outdoor weather measurements for
running the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline model fitting accuracy should have CV(RMSE) lower than 20%.
Thus, most research related to M&V for whole-building approach focuses
on the accuracy of baseline modeling, exploring model performance from
simple regression models to more complex machine learning techniques.
One study reviewed various models suitable for M&V applications as well
as selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.
become 'stale') over the extended period of pre- and post-analysis.

### Non-routine events impact

Non-routine events in M&V commonly refer to unexpected changes in a
building that influence a building's energy usage and are typically
unrelated to the intervention strategy. Current guidelines define
non-routine events as 'static factors' that needed to be adjusted after
projecting baseline in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies uses linear
interpolation to estimate counter-factual baseline for demand response
program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
one study points out that it is inaccurate to assume no change in the
operating conditions during the response period
[@huang_experimental_2023]. Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

@@

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
make it even simpler by adjusting the setpoint based on outdoor weather
conditions with more description shown in Section
\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
to:

@@

As a result, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption that
buildings are typically stable over the timeframe involved in
conventional M&V is rarely true. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events, especially operational drifts since the
assessment last for two years. And according to the statistics shown
here, it is a typical case observed in real buildings applications.

@@

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

@@

|                        |                                                                  |       |
|-------------------|------------------------------------|-------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

-   No carryover effect using a t-test with a p-value not exceeding a
    defined significance threshold of 0.05.

@@

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=12, fig.width=12}

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 30, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison of satisfying all stopping criteria") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison and for clarity, we plotted conventional M&V results in grey,
randomized M&V results after satisfying all stopping criteria in green,
and the randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Acc|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using TOWT model.
As a result, the conventional M&V method exhibits an average deviation
of 6% in savings estimation, whereas the proposed randomized method
demonstrates significantly smaller deviations. If the analyst stops
immediately after satisfying all stopping criteria, the deviation is
reduced to approximately 2%. Extending the M&V period to match the
length of the conventional method (24 months) further improves accuracy
to 1%. Furthermore, by comparing subplots (a) and (b), we observe a
minor improvement in estimation accuracy when a model is fitted to the
data to account for weather differences. The model fitting is most
beneficial for the shorter dataset (randomized and stopping when all
criteria are met, typically 24-36 weeks of data), where it improves
estimation accuracy by 0.6%. This is because the weather conditions
between the first and second year for each of the locations are
relatively similar on average over this time period, and thus the effect
of adjusting for temperature as an independent variable in the model is
relatively small. Subplot (c) compares the overall timeline of the two
M&V methods. The results show that almost all buildings achieve accurate
M&V results and meet the stopping criteria within 24 weeks (for this
intervention), with the remainder meeting all stopping criteria within
36 weeks. The additional 12 weeks is usually driven by a need to span
sufficient weather conditions representative of a full year of data.
Meanwhile, the red arrow highlights the excess time required by the
conventional M&V method (24 months). Thus, the randomized M&V method
finds more accurate results in approximately a quarter of the time
required for conventional M&V. Even if it was the case that an existing
12-month baseline data was available at the start of the M&V process,
randomized M&V would still obtain a more accurate result, in less time,
than obtaining the 12-month retrofit dataset required by conventional
M&V.

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events which are static changes or more gradual changes denoted as operational drift occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability. We also assessed the impact of starting the M&V at different time of the year, continuing randomization with different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.

@@

# colors
ls_colors <- c("Baseline" = "grey10",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#6baed6",
               "Randomized\n(50/50)" = "#6baed6",
               "Randomized\n(20/80)" = "#6baed6", 
               "Randomized\n(10/90)" = "#6baed6", 
               "Continued\n(20/80)" = "#6baed6", 
               "Continued\n(10/90)" = "#6baed6",
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Dropped (24 months)" = "#4292c6",
               "Kept (24 months)" = "#08519c",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. An early application of this
randomized design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance on two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defined stopping criteria to end the M&V period early
after reaching a target level of uncertainty, among other improvements,
such as using blocked randomization by weekday. The full framework is
detailed in a previously published study [@raftery2024] along with a
case study application to a real building, and other example use cases.
In summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially shortly after the test
begins. In addition, once the desired estimate uncertainty target is
achieved, analysts can terminate the M&V study and switch to 100%
intervention. The key advantage of randomization is that if control
strategies are sampled with equal probability, the influence of other
confounding factors such as occupancy change (commonly known as one of
the non-routine events) and other more subtle long term changes in
building energy use are likely to be evenly distributed among
measurements, leading to a more accurate and unbiased assessment of the
intervention effect.

@@

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly used the metered electricity and outdoor weather measurements for
running the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have CV(RMSE) lower than
25% for 12-month data collection. Thus, most research related to M&V for
whole-building approach focuses on the accuracy of baseline modeling,
exploring model performance from simple regression models to more
complex machine learning techniques. One study reviewed various models
suitable for M&V applications as well as selected input features
[@alrobaie_review_2022] and another study provided a definitive
methodology to apply machine learning models for M&V use cases
[@gallagher_development_2018]. In addition, a few studies investigated
the critical performance metrics to evaluate the developed baseline
models [@granderson_automated_2015; @granderson_development_2014] and
compared a variety of models using those metrics
[@granderson_accuracy_2016]. These studies made significant
contributions by emphasizing the uncertainty associated with the
model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.
become 'stale') over the extended period of pre- and post-analysis.

### Changes in building baseline measurements

Changes in a building that influences its energy usage are commonly
refered to as non-routine events in M&V and are typically unrelated to
the intervention strategy. Current guidelines define non-routine events
as 'static factors' that needed to be adjusted after projecting baseline
in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies uses linear
interpolation to estimate counter-factual baseline for demand response
program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
one study points out that it is inaccurate to assume no change in the
operating conditions during the response period
[@huang_experimental_2023]. Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

@@

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
made it even simpler by adjusting the setpoint based on outdoor weather
conditions with more description shown in Section
\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
to:

@@

As a result, only 66 buildings are labeled in this 'stable' subset (out
of total of 600 buildings) implying the assumption that buildings are
typically stable over the timeframe involved in conventional M&V is
rarely true. As mentioned in Section \@ref(non-routine-events-impact),
such variability is largely associated with non-routine events,
especially operational drifts since the assessment, even under ideal
situations, last for two years. And according to the statistics shown
here, it is a typical case observed in real buildings applications.

@@

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that resets the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

@@

|                        |                                                                  |       |
|-------------------|----------------------------------|-------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

@@

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 30, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison and for clarity, we plotted conventional M&V results in grey,
randomized M&V results after satisfying all stopping criteria in green,
and the randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Acc|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using TOWT model.
As a result, the conventional M&V method exhibits a median deviation of
5% in savings estimation, whereas the proposed randomized method
demonstrates significantly smaller deviations. If the analyst stops
immediately after satisfying all stopping criteria, the deviation is
reduced to approximately 1.6%. Extending the M&V period to match the
length of the conventional method (24 months) further improves accuracy
to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
observe a minor improvement in estimation accuracy when a model is
fitted to the data to account for weather differences. The model fitting
is most beneficial for the shorter dataset (randomized and stopping when
all criteria are met, typically 24-36 weeks of data), where it improves
estimation accuracy by 0.4%. This is because the weather conditions
between the first and second year for each of the locations are
relatively similar on average over this time period, and thus the effect
of adjusting for temperature as an independent variable in the model is
relatively small. Subplot (c) compares the overall timeline of the two
M&V methods. The results show that almost all buildings achieve accurate
M&V results and meet the stopping criteria within 24 weeks (for this
intervention), with the remainder meeting all stopping criteria within
36 weeks. The additional 12 weeks is usually driven by a need to span
sufficient weather conditions representative of a full year of data.
Meanwhile, the red arrow highlights the excess time required by the
conventional M&V method (24 months). Thus, the randomized M&V method
finds more accurate results in approximately a quarter of the time
required for conventional M&V. Even if it was the case that an existing
12-month baseline data was available at the start of the M&V process,
randomized M&V would still obtain a more accurate result, in less time,
than obtaining the 12-month retrofit dataset required by conventional
M&V.

We also plotted the estimated deviation without absolute calculation in
Figure \@ref(fig:meanerr) and we noticed that the uncertainty range
associated with the conventional method is significantly larger compared
to the randomized method that only takes 36 weeks according to subplot
(c) in Figure \@ref(fig:abstract).
